\documentclass{article}
\begin{document}
A section of text may be displayed by indenting it by an equal amount
on both sides, with the `quote' or `quotation' environment, as 
demonstrated here for the `quote' environment first.
\begin{quote}
Additonal vertical spacing is inserted above and below the displayed
text to seperate it vertically from normal text.

The text to be displayed may be over any length; it can be part of a
sentence, a whole paragraph, or several paragraphs.

Paragraphs are seperated as usual with an empty line, although no
empty lines are needed at the beginning and end of the displayed
text since additional vertical spacing is inserted here anyway.
\end{quote}
And here follows a demonstration of the `quotation' environment.
\begin{quotation}
In the `quotation' environment, paragraphs are marked by extra indentation
of the first line, wheras in the `quote' environment, they are indicated
more vertical spacing between them.

The present text is produced within the `quotation' environment, while
the example above was done with the `quote' environment.

The `quotation' environment is only really meaningfil, when the regular
text makes use of first-line indentation to show off new paragraphs.
\end{quotation}
The adjacent text behind the `quote' and/or `quotation' environment is
processed in standard form again, as shown here. An empty line behind
one of this environments starts the adjecent text as a new paragraph,
indicated by an indentend first line of the the following text. The
user may demonstrate this by the slight modification for this solution
example.
\end{document}




