\documentclass{article}
\begin{document}
A forfould nestin of the itemize an environment appears as follows:
\begin{itemize}
  \item The label for the first level is a black dot, a \emph{bullet}.
  \begin{itemize}
     \item That of the second level is a long dash (en dash).
     \begin{itemize}
        \item That of the third level is an asterisk.
        \begin{itemize}
           \item And the label for the fourth level is an upshifted dot.
           \item At the same time, the vertical spacing is decrease with
                 increasing depth.
        \end{itemize}  
        \item Back to the third level.
        \item Each level may contain several labels of the same degree,
        \item as shown here for third level.
     \end{itemize}
     \item Back to the second level.
  \end{itemize}
  \item And here we are at the first level of the itemize list again.
  \item The standard label markers for each level can be changed by
        the user
  \item as demonstrated with the solution example for exercise 4.6.
\end{itemize}
Similary for the enumerate environment, where the style of numberin changes
with the nesting level:
\begin{enumerate}
  \item The numbering at the first level is with Arabic numerals followed
        by a period.
  \begin{enumerate}
    \item At the second level, it is with lower case letters in paranthesis.
    \begin{enumerate}
      \item The third level is numbered with lower case Roman numerals
            with a period.
      \begin{enumerate}
        \item At the fourth level, capital letters are used.
        \item The label style can be changed as shown in the solution
              example for exercise 4.7.
      \end{enumerate}
      \item Back to the third level.
    \end{enumerate}
    \item Back to the second level.
    \item Also the enumerat environment may contain in every level
          several labels of the same degree,
    \item as shown here for the second level.
  \end{enumerate}
  \item Each level should contain at least two labels,
  \item but formally less labels are allowed!
\end{enumerate}                   
\end{document}





    
