\documentclass{article}
\begin{document}
\section*{PREFACE TO THE \LaTeX-GUIDE}

A new edition to ``A Guide to \LaTeX'' begs the fundamental question:
Has \LaTeX\ changed so much since the appearance of the third edition in 1999
that a new release of this manual is justified?

The simple answer to that question is `Well \dots.' In 1994, the \LaTeX\
world was in upheaval with the issue of the new version \LaTeXe, and the
second edition of the `Guide' came out just then to act as the bridge
between the old and new versions. By 1998, the initial teething problems had
been worked out and corrected through semi-annual releases, and the third
edition could describe an established, working system. However, homage was
still paid to the older 2.09 version since many users still employed its
familiar syntax, although they were most likely to be using it in a \LaTeXe\
environment. \LaTeX\ has now reached a degree of stability that since 2000
the regular updates have been reduced to annual events, which often appear
months after the nominal date, something that does not worry anyone. The old
version 2.09 is obsolete and should no longer play any role in such a manual.%
\footnote{With the result, that all maintance for \LaTeX\ 2.09 has been stopped
with the occurance of \LaTeXe\ as well as any further developement.}
In this fourth edition, it is reduced to an appendix just to document its
syntax and usage.

But if \LaTeX\ itself has not changed substantially since 1999, many of its
peripherals have. The rise of programs like pdf\TeX  and \texttt{dvipdfm} for
PDF output adds new possibilities, which are realized, not in \LaTeX\
directly, but by means of more modern \emph{packages} to extend the basic
features. The distribution of \TeX/\LaTeX\ installations has changed, such
that most users are given a complete, ready-to-run setup, with all the
`extras' that one used to have to obtain oneself. Those extras include
user-contributed packages, many of which are now considered indispensable.
Today `the \LaTeX\ system' includes much more than the basic kernel by Leslie
Lamport, encompassing the contributions of hundreds of other people. This
edition reflects this increase in breadth.\footnote{In between many hundreds
of supplements and extensions in form of packages have been developed for
\LaTeXe. Only the most important will be presented in this book.}

The changes to the fourth edition are mainly those of emphasis.
\begin{enumerate}
\item The material has been reorganized into `Basics' and `Beyond the Basics'
  (`advanced' sounds too intimidating) while the appendices contain
  topics that really can be skipped by most everyday users. Exception:
  Appendix H is an alphabetized command summary that many
  people find extremely useful (including ourselves).

  This reorganizing is meant to stress certain aspects over others. For
  example, the section on graphics inclusion and color was originally
  treated as an exotic freak, relegated to an appendix on extensions; in the
  third edition, it moved up to be included in a front chapter along with the
  \texttt{picture} environment and floats; now it dominates
  Chapter 6 all on its own, the floats come in the following
  Chapter 7, and \texttt{picturer} is banished to the later
  Chapter 13. This is not to say that the picture features are no good,
  but only that they are very specialized. We add
  descriptions of additional drawing possibilities there too.

\item It is stressed as much as possible that \LaTeX\ is a \emph{markup}
  language, with separation of content and form. Typographical settings
  should be placed in the preamble, while the body contains only logical
  markup. This is in keeping with the modern ideas of XML, where form and
  content are radically segregated.

\item Throughout this edition, contributed packages are explained at that point
  in the text where they are most relevant. The `fancyhdr' package
  comes in the section on page styles, \texttt{natbib} where literature
  citations are explained. This stresses that these `extensions' are part of
  the \LaTeX\ system as a whole. However, to remind the user that they must
  still be explicitly loaded, a marginal note is placed at the start of their
  descriptions.

\item PDF output is taken for granted throughout the book, in addition to the
  classical DVI format. This means that the added possibilities of `pdf\TeX'
  and `dvipdfm' are explained where they are relevant. A separate
  Chapter 10 on PostScript and PDF is still necessary, and the
  best interface to PDF output, the \texttt{hyperref package} by Sebastian
  Rahtz, is explained in detail. PDF is also included in
  Chapter 15 on presentation material.

  On the other hand, the other Web output formats, HTML and XML, are only
  dealt with briefly in Appendix E, since these are large topics
  treated in other books, most noticeably the \textsl{\LaTeX\ Web Companion}.

\item This book is being distributed with the \TeX Live CD, with the kind
  permission of Sebastian Rahtz who maintains it for the \TeX\ Users Group.
  It contains a full \TeX\ and \LaTeX\ installation for Windows, Macintosh,
  and Linux, plus many of the myriad extensions that exist.
\end{enumerate}

We once again express our hope that this \textsl{Guide} will prove more than
useful to all those who wish to find their way through the intricate world of
\LaTeX. And with the addition of the \TeX Live CD, that world is brought even
closer to their doorsteps.\footnote{Information for installation details are
given in the self documentation for the \TeX Live CD.}
\end{document}

