%%
%% This is file `AH_Bibliography.tex'
%% 
%% Copyright 2017 American Mathematical Society.
%% 
%% This file is part of the collection comprising the AMS Author Handbooks.
%% For details and license information, see the file README-AH.txt.
%%
%% The Current Maintainer of this work is the American Mathematical
%% Society.
%% 
%% ========================================================================
%% 

\renewcommand{\bibintro}{%
  Discounts are available on some of these books when they are ordered
  using information available on the \TeX\ Users Group page
  \href{https://www.tug.org/books/}{Books about TeX and Friends}.
  In particular, books published by Pearson affiliates (including
  Addison-Wesley) are eligible for a discount.}
  

\begin{thebibliography}{[ASMR]}

%\enlargethispage{2\baselineskip}
%\vspace*{-.75\baselineskip}
\raggedright

%\begin{extrabibtext}
%All documentation for AMS \tex/-related products is available in PDF
%form from the AMS web server as indicated below.  If you are reading
%this handbook on-line, the links for each item should be ``live''.
%\end{extrabibtext}

\bibitem[ABMR]{ABMR}
\href{http://www.ams.org/msnhtml/serials.pdf}
{\textit{Abbreviations of names of serials \textup[reviewed
in Mathematical Reviews\textup]}}, Amer. Math. Soc., Providence, RI.
\href{http://www.ams.org/msnhtml/serials.pdf}{\texttt{www.ams.org/msnhtml/serials.pdf}}

%\bibitem[AF]{AF} AMS Author FAQ,
%\url{www.ams.org/authors/author-faq}

\bibitem[AFG]{AFG} \href{ftp://ftp.ams.org/pub/tex/doc/amsfonts/amsfndoc.pdf}
{\textit{User's Guide to AMSFonts}}, version~2.2d,
Amer. Math. Soc.,  Providence, RI, 2002.  Link at
\href{http://www.ams.org/tex/amsfonts}{\texttt{www.ams.org/tex/amsfonts}}

\bibitem[AH]{AH} \textit{AMS Author Handbook},
Amer. Math. Soc., Providence, RI, 2014.  Link at
\href{http://www.ams.org/author-handbook}{\texttt{www.ams.org/author-handbook}}

\bibitem[AMG]{AMG} \href{http://mirrors.ctan.org/macros/latex/required/amsmath/amsldoc.pdf}{\textit{User's guide for the \pkg{amsmath} package
version~\textup{2.0}}}, Amer. Math. Soc., Providence, RI, and the \latex/3 Project, 2016.
\href{http://mirrors.ctan.org/macros/latex/required/amsmath/amsldoc.pdf}{\url{mirrors.ctan.org/macros/latex/required/amsmath/amsldoc.pdf}}
%\href{http://www.ams.org/tex/amslatex.html}{\texttt{www.ams.org/tex/amslatex}}

\bibitem[AMSR]{AMSR} \href{ftp://ftp.ams.org/pub/tex/amsrefs/amsrdoc.pdf}{\textit{User's Guide to the \pkg{amsrefs} Package}},
David Jones, Amer. Math. Soc., Providence, RI, 2013. Link at 
\href{http://www.ams.org/authors/amsrefs}{\texttt{www.ams.org/authors/amsrefs}}

%\bibitem[ATG]{ATG}
%\href{ftp://ftp.ams.org/pub/tex/doc/amstex/amsguide.pdf}
%{\textit{User's Guide to \amstex/}}, version~2.2,
%Amer. Math. Soc., Providence, RI, 2001.
%Link at \url{www.ams.org/tex/amstex}

\bibitem[ATH]{ATH}
\href{ftp://ftp.ams.org/pub/tex/doc/amscls/amsthdoc.pdf}
{\textit{Using the \texttt{amsthm} Package}},
version~2.20.2, Amer. Math. Soc., Providence, RI, 2015.  Link at 
\href{http://www.ams.org/tex/amslatex.html}{\texttt{www.ams.org/tex/amslatex}}

%\bibitem[ATP]{ATP} \textit{Using the \pkg{amsthm} package,
%version~\textup{2.0}}, Amer. Math. Soc., Providence, RI, 2004.\\
%Link at \url{www.ams.org/tex/amslatex.html}

\ifproceedings
\bibitem[EDP]{EDP}
\href{http://www.ams.org/publications/editpkg}
{Guide to AMS Editor's Package},
\url{www.ams.org/publications/editpkg}.
\fi

\bibitem[FAQ]{FAQ}
\href{http://www.ams.org/authors/author-faq}
{\textit{Frequently Asked Questions for AMS Authors}},
\url{www.ams.org/authors/author-faq}.
%%\url{www.ams.org/author-faq}.

%\bibitem[INT]{INT}
%\href{ftp://ftp.ams.org/pub/tex/doc/amstex/instr-t.pdf}
%{\textit{Instructions for Preparation of Papers and
%Monographs\textup{:} \amstex/}}, version~2.2b,
%Amer. Math. Soc., Providence, RI, 2002.
%Link at \url{www.ams.org/tex/amstex}

%\begin{extrabibtext}
%An extensive, up-to-date list of \tex/-related publications (both in print
%and electronic) appears on the web page
%\url{www.ams.org/tex/tex-pub} with links to sources where
%these publications can be obtained.
%\end{extrabibtext}

\bibitem[GM]{GM}  Michel Goossens, Frank Mittelbach et al.,
\textit{The \latex/ Graphics companion}, second ed., Addison-Wesley, Reading,
MA, 2007.

\bibitem[GGr]{GGr} George Gr\"atzer, \textit{Math into \latex/}, third ed.,
Springer, New York, 2000.

%\bibitem[Gr]{Gr} George Gr\"atzer, \textit{More Math into \latex/}, fourth ed.,
%Springer, New York, 2007.  (Supersedes \cite{GGr}.)

\bibitem[Gr]{Gr} George Gr\"atzer, \textit{More Math into \latex/}, fifth ed.,
Springer, New York, 2016.  (Supersedes \cite{GGr}.)

\bibitem[Gr1]{Gr1}
G.~Gr{\"a}tzer.
\newblock What is new in \latex/? I. Breaking free.
\newblock {\em Notices of the AMS}, 56(1):52--54, January 2009.
\href{http://www.ams.org/notices/200901/tx090100052p.pdf}{\texttt{www.ams.org/notices/200901/tx090100052p.pdf}}

\bibitem[Gr2]{Gr2}
G.~Gr{\"a}tzer.
\newblock What is new in \latex/? II. \tex/ implementations, evolution or
  revolution.
\newblock {\em Notices of the AMS}, 56(5):627--629, May 2009.
\href{http://www.ams.org/notices/200905/rtx090500627p.pdf}{\texttt{www.ams.org/notices/200905/rtx090500627p.pdf}}

\bibitem[Gr3]{Gr3}
G.~Gr{\"a}tzer.
\newblock What is new in \latex/? III. Formatting references.
\newblock {\em Notices of the AMS}, 56(8):954--956, September 2009.
\href{http://www.ams.org/notices/200908/rtx090800954p.pdf}{\texttt{www.ams.org/notices/200908/rtx090800954p.pdf}}

\bibitem[Gr4]{Gr4}
G.~Gr{\"a}tzer.
\newblock What is new in \latex/? IV. WYSIWYG \latex/.
\newblock {\em Notices of the AMS}, 58(6):828--830, June/July 2011.
\href{http://www.ams.org/notices/201106/rtx110600828p.pdf}{\texttt{www.ams.org/notices/201106/rtx110600828p.pdf}}

\bibitem[Gr5]{Gr5}
G.~Gr{\"a}tzer.
\newblock What is new in \latex/? V. \latex/ on an iPad. Foundation.
\newblock {\em Notices of the AMS}, 60(3):332--334, March 2013.
\href{http://www.ams.org/notices/201303/rnoti-p332.pdf}{\texttt{www.ams.org/notices/201303/rnoti-p332.pdf}}

\bibitem[Gr6]{Gr6}
G.~Gr{\"a}tzer.
\newblock What is new in \latex/? VI. \latex/ on an iPad. Empire.
\newblock {\em Notices of the AMS}, 60(4):434--439, April 2013.
\href{http://www.ams.org/notices/201304/rnoti-p434.pdf}{\texttt{www.ams.org/notices/201304/rnoti-p434.pdf}}

\bibitem[Gr7]{Gr7}
G.~Gr{\"a}tzer.
\newblock What is new in \latex/? VII. The STIX math symbols.
\newblock {\em Notices of the AMS}, 62(6):667--669, June/July 2015.
\href{http://www.ams.org/notices/201506/rnoti-p667.pdf}{\texttt{www.ams.org/notices/201506/rnoti-p667.pdf}}

\bibitem[Hi]{Hi} Nicholas J. Higham, \textit{Handbook of Writing for
the Mathematical Sciences}, second ed., SIAM, Philadelphia, PA, 1998.

\bibitem[KD]{KD} Helmut Kopka and Patrick W. Daly, \textit{Guide to \latex/},
fourth ed., Addison-Wesley, Boston, 2004.

\bibitem[Kn]{Kn} Donald E. Knuth, \textit{The \TeX book},
Addison-Wesley, Reading, MA, 1984.

\bibitem[La]{La} Leslie Lamport, \textit{\LaTeX{\textup:} A document
preparation system}, second ed., Addison-Wesley, Reading, MA, 1994.

\bibitem[LW]{LW} Mary Letourneau and Jennifer Wright Sharp,
\textit{AMS style guide: Journals}, Amer. Math. Soc., Providence, RI, 2017.
\href{http://www.ams.org/publications/authors/AMS-StyleGuide-print.pdf}{\texttt{www.ams.org/publications/authors/AMS-StyleGuide-print.pdf}}

\bibitem[MeSlp]{MeSlp}
Andrew Mertz, William Slough,
\emph{Graphics with \TikZ},
Prac\tex/ Journal, \textbf{2007}:1.
\href{http://tug.org/pracjourn/2007-1/mertz/mertz.pdf}{\texttt{tug.org/pracjourn/2007-1/mertz/mertz.pdf}}

\bibitem[MeSlt]{MeSlt}
Andrew Mertz, William Slough,
\emph{Graphics with \textsc{pgf} and \TikZ},
TUGboat, \textbf{28}:1 (2007), 91--99.
\href{http://tug.org/TUGboat/tb28-1/tb88mertz.pdf}{\texttt{tug.org/TUGboat/tb28-1/tb88mertz.pdf}}

\bibitem[MG]{MG} Frank Mittelbach, Michel Goossens, et al.,
\textit{The \latex/ companion}, second ed., Addison-Wesley, Reading,
MA, 2004.  This is now also available as an ebook, in both English and
German; see the TUG web page cited above.  The front matter, including
the full Table of Contents, can be viewed online, at
\href{https://www.latex-project.org/help/books/tlc2-ch0.pdf}{\texttt{www.latex-project.org/help/books/tlc2-ch0.pdf}}.

\bibitem[MSC]{MSC}
2010 Mathematics Subject Classification
\href{http://www.ams.org/msc}{\texttt{www.ams.org/msc}}.

\bibitem[SHSD]{SHSD}
Norman E. Steenrod, Paul R. Halmos, Menahem M. Schiffer, and Jean A.
Dieudonn\'e, \textit{How to write mathematics}, 6th printing (2000),
Amer. Math. Soc., Providence, RI, 1973, reprinted with corrections 1981.

\bibitem[Sw]{Sw} Ellen E. Swanson, 
%\href{ftp://ftp.ams.org/pub/author-info/documentation/howto/mit-2.pdf}
\href{http://www.ams.org/publications/authors/mit-2.pdf}
{\textit{Mathematics into type}},
updated ed., Amer. Math. Soc., Providence, RI, 1999.  Link at
\href{http://www.ams.org/publications/authors}{\texttt{www.ams.org/authors}}

\bibitem[tkz]{tkz}
Till Tantau,
The \TikZ\ and PGF Packages: Manual for version 3.0.1a, 
2013.
\href{http://www.ctan.org/tex-archive/graphics/pgf/base/doc/pgfmanual.pdf}{\texttt{www.ctan.org/tex-archive/graphics/pgf/base/doc/pgfmanual.pdf}}

\bibitem[tkz-ex]{tkz-ex}
Online gallery of \TikZ\ and PGF examples.
\href{http://www.texample.net/tikz/examples/all/}{\texttt{www.texample.net/tikz/examples/all/}}

\bibitem[voss-pst]{voss-pst}
Herbert Voss,
PSTricks: Graphics and PostScript for \TeX\ and \LaTeX,
UIT Cambridge Ltd.,
2011.

\bibitem[voss-tbl]{voss-tbl}
Herbert Voss,
Typesetting Tables with \LaTeX,
UIT Cambridge Ltd.,
2010.

\end{thebibliography}

\endinput
