%%
%% This is file `Submitting2AMS.tex'
%% 
%% Copyright 2017 American Mathematical Society.
%% 
%% This file is part of the collection comprising the AMS Author Handbooks.
%% For details and license information, see the file README-AH.txt.
%%
%% The Current Maintainer of this work is the American Mathematical
%% Society.
%% 
%% ========================================================================
%% 

\chapter{Submitting files to the AMS}\label{ch:submit}

\section{Submission guidelines}

\noindent Upon acceptance of your 
\jmpm{article}{book}{paper}{\Memo}, the source file(s) should be sent
to the AMS office (this includes the \tex/ 
source file(s) and any graphics files).  Send
\emph{only} the files that are needed to process your submission or archive
it for future reference.  (This does \emph{not} include \fn{.log} or
\fn{.aux} files, for example.)

Before sending the source file(s), be sure you have proofread your 
\jmpm{article}{monograph}{paper}{\Memo\ submission} carefully. 
The files that you send must be the EXACT files
used to generate the proof copy that was accepted for publication.  In
order to avoid any possible production problems, before sending the files,
be sure to verify all items in the sections \textit{\nameref{sec:check}} (page~\pageref{sec:check}) and \textit{\nameref{sec:scheck}} (page~\pageref{sec:scheck}).

If your submission consists of \textbf{multiple files}, we recommend that
you bundle them using the Zip utility; this can be obtained (free) for
most platforms from \href{http://freecode.com}{\texttt{freecode.com}}\,.
Bundling means that only one (compressed) file needs to be sent, lessening
the chance of name conflicts or file corruption.

\section{Web server submissions (preferred)}
\noindent Accepted electronic manuscripts can be submitted via the web server
at\\
\href{http://www.ams.org/submit-book-journal}{\texttt{www.ams.org/submit-book-journal}}.
For security and confidentiality
reasons, submitting through the web server requires an AMS web account.
Authors who do not already have an account will be given the opportunity
to create one as they go through the submission process.

\section{Electronic mail submissions}
\noindent
Files sent by electronic mail should be addressed to
\texttt{pub-submit@ams.org}.  Include them as attachments, not as
part of the message.

The subject line of the message should use the publication code to identify the
\jmpm{journal (see the list of \textit{\nameref{tbl:series}}, page~\pageref{tbl:series}).}
{monograph series (see the list of \textit{\nameref{tbl:series}},
 page~\pageref{tbl:series}), and should include the name(s) of the author(s).}
{proceedings/collection series (see the list of \textit{\nameref{tbl:series}},
 page~\pageref{tbl:series}), and should include the name(s) of the editor(s).}
{as a \Memos\ submission.}
By including this information in the subject line, you will help speed up
the processing of your submission.

Submissions received through email will be acknowledged upon receipt by an
automatic reply while your submission is reviewed.  If there are any problems
with the file received, you will be notified.

\section{Other possibilities}
\noindent If your attempt to submit both through the web server and by
electronic mail fails, arrangements can be made for you to post your
files via FTP or on physical media.  Requests for help can be addressed
as described in the section ``\nameref{sec:amsresources}'' on
page~\pageref{sec:amsresources}.

\endinput
