\documentclass{article}
\begin{document}
\begin{thebibliography}{\hspace{1.5cm}}
\bibitem[abr90]{abr90} Abraham P. W., with Hargraves K. A. and Berry K. (1990).
   \textsl{\TeX\ for the Impatient.} Reading MA: Addison-Wesley
\bibitem[bec93]{bec93} Bechtolsheim, S. von (1993) \textsl{\TeX\ in Practice,
   Vol.\ I--IV,} New York: Springer
\bibitem[bec93a]{bec93a} Vol.\ I: Basics
\bibitem[bec93b]{bec93b} Vol.\ II: Paragraphs, Math, and Fonts
\bibitem[bec93c]{bec93c} Vol.\ III: Tokens, Macros
\bibitem[bec93d]{bec93d} Vol.\ IV: Output Routines, Tables
\bibitem[dal92]{dal92} Daly P. W. and Kopka H. (1992) {A Guide to \LaTeX.}
   Workingham, UK: Addison-Wesley
\bibitem[dal95]{dal95} Daly P. W. and Kopka H. (1995) {A Guide to \LaTeX,}
   2nd ed. Workingham, UK: Addison-Wesley
\bibitem[dal99]{dal99} Daly P. W. and Kopka H. (1999) {A Guide to \LaTeX,}
   3rd ed. Workingham, UK: Addison-Wesley
\bibitem[dal03]{dal03} Daly P. W. and Kopka H. (2003) {A Guide to \LaTeX,}
   4th ed. Workingham, UK: Addison-Wesley
\bibitem[eij92]{eij92} Eijkhout V. (1992). \textsl{\TeX\ by Topic, a 
   \TeX nician's Referenc.} Harlow: Addison-Wesley
\bibitem[goo94]{goo94} Goossens M., Mittelbach F. and Samarin A. (1994).
   \textsl{The \LaTeX\ Companion.} Reading MA: Adddison-Wesley
\bibitem[goo97]{goo97} Goossens M., Rahtz S. and Mittelbach F. (1997).
   \textsl{The \LaTeX\ Graphic Companion.} Reading MA: Addison-Wesley
\bibitem[goo99]{goo99} Goossens M. and Rahtz S. (1999).
   \textsl{The \LaTeX\ Web Companion.} Reading MA: Addison-Wesley
\bibitem[knu91]{knu91} Knuth D. E. (1991) \textsl{Computers and Typesetting
   Vol A--E.}  Reading, MA: Addison-Wesley
\bibitem[knu91a]{knu91a} Knuth D. E. (1991a) Vol. A: \textsl{The \TeX book.}
   11.~ed. Reading MA: Addison-Wesley
\bibitem[knu91b]{knu91b} Knuth D. E. (1991b) Vol. B: \textsl{The Program.}
   4.~ed. Reading MA: Addison-Wesley
\bibitem[lam94]{lam94} Lamport L. (1994) \textsl{\LaTeX---A Document
   Preparation System}, 2nd ed. Reading MA: Addison-Wesley
    Addison-Wesley-Longman (Deutschland) GmbH. Bonn 1995
\end{thebibliography}
\end{document}
