\documentclass[a4paper]{article}
\usepackage[talos]{fontsetup}

\pagestyle{empty}
\begin{document}

\begin{center}
  \textbf{\Large The Talos fonts}\\
v1.0 31/Jan/2022\\[1ex]
Antonis Tsolomitis\\
University of the Aegean,
  Department of Mathematics\\
  832\,00 Karlovassi, Samos, Greece
\end{center}

The Talos fonts were bitmap fonts created especially for
a custom installation in a VMS machine bought in the eighties
by the Department of Mathematics of the University of Crete, Greece.
This was definitely the first TeX installation in a Greek University
and most probably the first TeX installation that supported the Greek
(monotonic) language.

The VMS server, with hostname ``Talos'', after the name of the Cretan mythical
robot, made by Hephaestus to protect Europa at the request of Zeus,
was connected with dummy terminals to the professors' offices. There was no
previewer other than printing the postscript file.
The server was in use in the department of Mathematics until \textsc{1998}
when it suffered a severe crash and was abandoned. In that machine,
the ``Talos'' (Τάλως in Greek), many professor notes, student thesis and
dissertations were written
with its special Greek font that we now revive. The characters were redrawn
by some printouts that I had carefully saved for this purpose, when I was
employed in the late nineties in the department. I am positive that
many colleagues still have printouts from that era and the Talos server.

The Latin characters (including mathematics) were cmr and in this release
are drawn from NewComputerModern Book weight. However, the Greek part
is of unknown origin. The design is very interesting especially the italic one.
The Italic capitals use what we know as ``mathcal'' for the characters
that are common with Latin and Greek but have matching custom designs
for non common ones.

The Italic Greek characters will soon find their way into NewComputerModern Math fonts as
a Greek calligraphic.

Lowercase Greek:
\begin{center}
  αβγδεζηθικλμνξοπρσςτυφχψω άέήίόύώ ϊϋ ΐΰ\\
  \textbf{αβγδεζηθικλμνξοπρσςτυφχψω άέήίόύώ ϊϋ ΐ\ ΰ}
\end{center}

Lowercase Greek Italic:
\begin{center}
  \textit{αβγδεζηθικλμνξοπρσςτυφχψω άέήίόύώ ϊϋ ΐ\ ΰ}\\
   \textit{\bfseries αβγδεζηθικλμνξοπρσςτυφχψω άέήίόύώ ϊϋ ΐ\ ΰ}
\end{center}

Uppercase Greek:
\begin{center}
  ΑΒΓΔΕΖΗΘΙΚΛΜΝΞΟΠΡΣΣΤΥΦΧΨΩ ΆΈΉΊΌΎΏ ΪΫ\\
  \textbf{ΑΒΓΔΕΖΗΘΙΚΛΜΝΞΟΠΡΣΣΤΥΦΧΨΩ ΆΈΉΊΌΎΏ Ϊ\ Ϋ}
\end{center}

Uppercase Greek Italic:
\begin{center}
  \textit{Α\,Β\,Γ\,Δ\,Ε\,Ζ\,Η\,Θ\,Ι\,Κ\,Λ\,Μ\,Ν\,Ξ\,Ο\,Π\,Ρ\,Σ\,Τ\,Υ\,Φ\,Χ\,Ψ\,Ω\,Ά\,Έ\,Ή\,Ί\,Ό\,Ύ\,Ώ\,Ϊ\,Ϋ}\\
   \textit{\bfseries Α\,Β\,Γ\,Δ\,Ε\,Ζ\,Η\,Θ\,Ι\,Κ\,Λ\,Μ\,Ν\,Ξ\,Ο\,Π\,Ρ\,Σ\,Τ\,Υ\,Φ\,Χ\,Ψ\,Ω\,Ά\,Έ\,Ή\,Ί\,Ό\,Ύ\,Ώ\,Ϊ\,Ϋ}
\end{center}
%\enlargethispage{2cm}
A sample with mathematics (from NewCM) can be found in the \verb|fontsetup| documentation.
Load the fonts with \verb|\usepackage[talos]{fontsetup}|

Metrics may change in future releases especially kerning pairs.



\end{document}
